---
title: "My Kind of Movies App"
author: "Chris van Hasselt"
date: "July 2, 2016"
output: 
  ioslides_presentation:
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## My Kind of Movies App
The "My Kind of Movies!" app is a simple Shiny app based on data found through the
[Grouplens Project](http://grouplens.org/datasets/movielens/)
and usese data from their [benchmark dataset](http://grouplens.org/datasets/movielens/100k/)
of 100,000 movie ratings of 1,700 different movies from one-thousand survey respondents.

The app effectively demonstrates some of the things that Shiny can do.  While
barely touching on Shiny's full capabilities, the app shows how a complicated 
set of data can be quickly "crunched" to provide value from data.

## Approach to Data
Preparing the data from Grouplens for use in the app involved a number of steps: 

- Renaming column variables for convenience
- Changing the data types on some columns, particularly date and logical columns
- Defining an age group column, rather than just using user age
- Saving the modified data files

The logical columns, for movie genre, were  ultimately not used.
The file _dataTidying.R_, found on the [GitHub repository](https://github.com/cvanhasselt/CvH_DataProduct.git), includes code
used to prepare data for the app.

- The data includes information on movies, user demographics, and movie ratings.  
- All three of these data sets were combined;
- The user filters the data by age group, gender and occupation.
- The most difficult task was determining an effective way to rank the movies by rating, given
that not all movies had the same number people rating them
--A movie with 19 users giving the movie a 5, and one user giving it a 4 would have 
a lower rank than a movie with one user rating of 5.


## Ranking Algorithm - Part 1
The ranking algorithm used follows the formula
found [here on Stack Overflow](http://stackoverflow.com/a/1411268/941255).  

An example of this ranking approach is demonstrated below.  The data frame does not correspond
to the data frame format used in the application. 
```{r rankdemo, include=FALSE}

StarWars <- c(movie.id=50,title="Star Wars (1977)", avgRating=4.533333333,numRatings=15.0)

RagingBull <-c(movie.id=192,title="Raging Bull (1980)", avgRating=4.75,numRatings=4)
movieDF <- data.frame(rbind(StarWars,RagingBull))
```
```{r demoparta, echo=FALSE}

knitr::kable(movieDF)

```

## Ranking Algorithm - Part 2
```{r demopartb, echo=TRUE}
v1 <- as.numeric(as.character(movieDF["StarWars",]$numRatings))
R1 <- as.numeric(as.character(movieDF["StarWars",]$avgRating))
v2 <- as.numeric(as.character(movieDF["RagingBull",]$numRatings))
R2 <- as.numeric(as.character(movieDF["RagingBull",]$avgRating))
m <- 3 # arbitrary choice of a minimum number of items
# I've hardcoded this, but in practice it is calculated across a large set of movie ratings
C = 3.245463925 

ScoreStarWars <- (v1 /(v1 + m)) * R1 + (m/(v1 + m)) * C
ScoreRagingBull <- (v2 /(v2 + m)) * R2 + (m/(v2 + m)) * C
# Star Wars Score
ScoreStarWars
# Raging Bull Score
ScoreRagingBull
```
As can be seen, final score is better for _Star Wars_, even though _Raging Bull_ 
had a higher raw score.

## Conclusion

The _My Kind of Movies_ app demonstrates the power of Shiny.  

There is a lot of room for improvement, but the app demonstrates the power of bringing
data to life through an effective, interactive presentation.

